---
title: "An introduction to `rlogging`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{An introduction to `rlogging`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
**rlogging** is an helpful package providing an easy way to setup a logging system in R.

## Setup
Firstly, we load `rlogging`:

```{r setup}
library(rlogging)
```

## rlogging
We present here a simple example to show how to use `rlogging`.

### A `Logger`
The `?Logger` class is the core of the logging system. 

A `Logger` object is a `list` consisting of 4 elements: 

* `path`: (optional) character vector, a path to the log file to use
* `con`: (optional) connection to a file
* `verbose`: logical, whether R should report extra information on progress
* `level`: character vector, the log level of the logger.

Eight levels are supported:

* `"OFF"`: indicates the logger is inactive
* `"FATAL"`: indicates a fatal event in the application that prevents it from running
* `"ERROR"`: indicates an error in the application, possibly recoverable
* `"WARN"`: indicates an event that might lead to an error
* `"INFO"`: indicates informational messages
* `"DEBUG"`: indicates general debugging messages
* `"TRACE"`: indicates fine-grained debug events
* `"ALL"`: all levels are considered

Levels are considered ordered: `OFF < FATAL < ERROR < WARN < INFO < DEBUG < TRACE < ALL`.
For example, if we set the level of a logger to `DEBUG`, then all `INFO`, `WARN`,
`ERROR`, and `FATAL` events will be logged.

Let's create some loggers with different levels:

```{r logger_setup}
#Logger with `level` set to "INFO"
loggerInfo  = createLogger(level = "INFO")
#Logger with `level` set to "DEBUG"
loggerDebug = createLogger(level = "DEBUG")
#Logger with `level` set to "TRACE"
loggerTrace = createLogger(level = "TRACE")
```

### Example
Let's create a simple function to show how to use `rlogging`. For example, we 
can write a function to calculate the sum of two numbers `a` and `b`:

```{r function_setup}
mySumFunction <- function(
    a = 1,
    b = 2,
    logger
  ){
  #We insert here an informational message
  logInfo(object = logger, message = "Start of mySumFunction", sep = "\n", add.level = T, add.time = T)
  
  #We insert here a fine-grained informational message (trace level)
  logTrace(object = logger, message = "Check if logger has a connection to a file and (if so) open it...", add.level = T, add.time = T)
  logger = openCon(object = logger)
  logTrace(object = logger, message = "DONE", sep = "\n")
  
  #We insert here a fine-grained informational message (debug level)
  logDebug(object = logger, message = "Check input...", add.level = T, add.time = T)
  if(isFALSE(is.numeric(a))){stop("'a' must be numeric")}
  if(isFALSE(is.numeric(b))){stop("'b' must be numeric")}
  logDebug(object = logger, message = "DONE", sep = "\n")
  
  #We insert here a fine-grained informational message (debug level)
  logDebug(object = logger, message = "Compute sum...", add.level = T, add.time = T)
  out = a + b
  logDebug(object = logger, message = "DONE", sep = "\n")
  
  #We insert here a fine-grained informational message (trace level)
  logTrace(object = logger, message = "Close an opened connection...", add.level = T, add.time = T)
  closeCon(object = logger)
  logTrace(object = logger, message = "DONE", sep = "\n")
  
  #We insert here an informational message
  logInfo(object = logger, message = "End of mySumFunction", sep = "\n", add.level = T, add.time = T)
  
  return(out)
}
```

The defined function has three arguments in input:

* `a`: a numeric value
* `b`: a numeric value
* `logger`: an object of class `Logger`

In the function above, we inserted various messages with different logging
levels. Let's see what happens when we use different loggers.

For example, let's compute the sum `3 + 4`.

### Log level: `INFO`
Let's check what happens if we use a logger with the logging level set to 
`"INFO"`.

```{r log_level_info}
mySumFunction(
  a = 3,
  b = 4,
  logger = loggerInfo
)
```

As expected, only the `INFO` messages are reported.

### Log level: `DEBUG`
Let's check what happens if we use a logger with the logging level set to 
`"DEBUG"`.

```{r log_level_debug}
mySumFunction(
  a = 3,
  b = 4,
  logger = loggerDebug
)
```
As expected, all the messages with level `DEBUG` and above (i.e., 
`INFO`) are reported.

### Log level: `TRACE`
Let's check what happens if we use a logger with the logging level set to 
`"TRACE"`.

```{r log_level_trace}
mySumFunction(
  a = 3,
  b = 4,
  logger = loggerTrace
)
```

As expected, all the messages with level `TRACE` and above (i.e., `DEBUG` and
`INFO`) are reported.
